---
title: "ANA600 Final Exam"
author: "Charlene Wolthers"
date: "`r Sys.Date()`"
output: html_document
---
## INSTRUCTIONS 

Perform basic exploratory data analysis (EDA) for the final exam. EDA consists of the procedures and concepts we have practiced throughout the course. You will be exploring your dataset, reviewing the variables, modeling an variable of interest, and interpreting results. Each section requires a complete paragraph explaining and interpreting the results within the R-Markdown above the code block required to analyze that section. All writing and code should be written as if presenting a report to your supervisor.

Provided data was selected to gather information about consumer attitudes and decisions to save, borrow, or make purchases. The researchers wanted to forecast changes in consumer behavior in the United States and better understand consumer confidence about personal finances, employment, price changes, and the perceived state of national business. Your job is to develop a model of income by sex to determine if there is a statistically significant difference between the two.

Below is a list of the variables included in the dataset:

* household: including yourself, how many members of your household are 18 years or older?
* kids: How many members of your household are 17 years or younger?
* vehicles: How many vehicles do you use, including those leased, owned, or provided by your employer for personal use?
* priceExpected: By what percent do you expect prices to go up/down on average during the next 12 months?
* incomeExpected: By what percent per year do you expect your (family) income to increase/decrease during the next 12 months?
* businessExpected: Considering business conditions in the country as a whole, do you think that during the next 12 months, we'll have good or bad times financially?
* financialStability: Would you say that you (and your family) are better off or worse off financially than you were a year ago?
* investments: What type of investments do you plan to add or shift money into or open during the next three months: mutual funds, savings accounts, stocks, bonds, retirement accounts?
* income: What is your household gross income (in thousands)?
* age: What is your age?
* employmentSector: What is your employment sector?
* region: What is your state of residence? (recoded into regions)
* hoursPerWeek: Approximately how many hours a week are you employed?
  
***
***

## PREPARATION (10 Points)
*Import the required file to a new dataframe and load necessary libraries*

```{r include=FALSE}

myDataLocation <- "C:/Users/rnehe/Desktop/ANA600/Wk4/"
setwd(myDataLocation)
acsData <- read.csv(file = "ACS.csv", header = TRUE)

library(dplyr)
library(ggplot2)
library(ggformula)
library(mosaic)

```

## INTRODUCTION AND RESEARCH QUESTION (5 Points)
1. Introduce your purpose and scope, creating a story for the data generation process that might be responsible for the variation in income output variable.
2. Describe the research question, which is to examine income based on full-time or part-time work hours per week.

***

## QUESTION #1 (10 Points)
1. Enter code to produce the structure of your dataframe
2. Recode the Sex variable to 0=Female, 1=Male
3. Produce a crosstab table of observations for the race and sex variables 
4. View the top five records in the dataframe
5. Write one paragraph describing the structure of the data frame and interpreting the produced table

{Your writeup here}

```{r Question_01}

str(acsData)

acsData$Sex <- factor(
  acsData$Sex,
  levels = c(0, 1),
  labels = c("Female", "Male")
)

tally(Race ~ Sex, data = acsData)



```

## QUESTION #2 (10 Points)
1. Recode the income variable to value x 1,000
2. Calculate the minimum, maximum, mean, median, IQR, and range for income
3. Calculate the mean of income each for males and for females
3. Write one paragraph explaining and interpreting the descriptive statistics

{Your writeup here}

```{r Question_02}
acsData <- read.csv(file = "ACS.csv", header = TRUE) ###DELETEME
acsData$Income <- acsData$Income * 1000

acsData.stats <- favstats(acsData$Income)
acsData.mean <- acsData.stats$mean
acsData.stats

acsData.Females <- filter(
  acsData,
  Sex == 0
)
acsData.meanF <- mean(acsData.Females$Income)

acsData.Males <- filter(
  acsData,
  Sex == 1
)
acsData.meanM <- mean(acsData.Males$Income)

acsData.mean
acsData.meanF
acsData.meanM

```

## QUESTION 3 (10 Points)
1. Create a new variable SexL such that Sex = 0 is "Female", Sex = 1 is "Male", and else is "Undefined" 
2. Create an appropriate visualization for income and the new SexL variable
3. Write one paragraph explaining and interpreting the visualization

{Your writeup here}

```{r Question_03, echo=FALSE}

acsData <- acsData %>%
  mutate(SexL = case_when(
    Sex == 0  ~ "Female",
    Sex == 1  ~ "Male",
    TRUE     ~ "Undefined"
  ))

ggplot(
  acsData, 
  aes(
    x = SexL, 
    y = Income
    )
  ) +
  geom_boxplot(
    fill = "lightblue", 
    color = "blue", 
    width = 0.7
    ) +
  scale_y_continuous(
    labels = dollar_format(
      scale = 1, prefix = "$"
      )
    ) +
  labs(
    title = "Comparative Boxplot of Income by Sex", 
    x = "Sex", 
    y = "Income"
    )

```

## QUESTION 4 (10 Points)
1. Create a histogram of income by sex with facet grid
3. Write one paragraph explaining and interpreting the visualization

{Your writeup here}

```{r Question04, echo=FALSE}

gf_histogram(
  ~Income, 
  data=acsData, 
  fill=~SexL, 
  bins=20
  ) |>
  gf_labs(
    title = "Distribution of Age by Gender", 
    x="Income") |>
  gf_facet_grid(SexL ~.)

```
## QUESTION 5 (10 Points)
1. Create a model of income for females
3. Write one paragraph explaining and interpreting the model

{Your writeup here}

```{r Question05, echo=FALSE}

acsData.modelF <- lm(Income ~ NULL, data = acsData)
acsData.modelF

```

## QUESTION 6 (20 Points)
1. Create a model of income by sex
3. Write one paragraph explaining and interpreting the model

{Your writeup here}

```{r Question06, echo=FALSE}

acsData.modelM <- lm(Income ~ SexL, data = acsData)
acsData.modelM

```

## QUESTION 7 (20 Points)
1. Calculate the predicted value and residual value for each observation using the income by sex model
2. Calculate the sum of squared deviations and sum of absolute deviations
3. Write one paragraph explaining and interpreting the results

{Your writeup here}

```{r Question_07}

acsData$Prediction <- predict(acsData.modelM)
acsData$Residual <- resid(acsData.modelM)

acsData.modelM.SSD <- sum(acsData$Residual^2)
acsData.modelM.SAD <- sum(abs(acsData$Residual))

acsData.modelM.SSD
acsData.modelM.SAD

```

## QUESTION 8 (20 Points)
1. Run an analysis of variance on the model of income by sex
2. Write one paragraph explaining and interpreting the results
Hint:
* H0: bi = 0
* Ha: bi <> 0

{Your writeup here}

```{r Question_08}

acsData.anovaF <- anova(acsData.modelF)
acsData.anovaF

acsData.anovaM <- anova(acsData.modelM)
acsData.anovaM

```

## QUESTION 9 (10 Points)
1. Calculate the proportional reduction in error
2. Write one paragraph explaining and interpreting the improvement of the linear model by adding sex
Hint:
* H0: bi = 0
* Ha: bi <> 0

{Your writeup here}

```{r Question_09}

acsData.PRE_F <- acsData.anovaF$`Sum Sq`[1]
acsData.PRE_F

acsData.PRE_M <- acsData.anovaM$`Sum Sq`[2]
acsData.PRE_M

acsData.PRE <- (acsData.PRE_F - acsData.PRE_M) / acsData.PRE_F
acsData.PRE

#acsData.CD <- cohensD(Income ~ Sex, data = acsData)
#acsData.CD

```

## QUESTION 10 (20 Points)
Write one paragraph interpreting and concluding the results of your analysis.

{Your writeup here}

**END OF FINAL EXAM**